# Other figures

## Ecosystem fidelity
All indicators are assigned to at least one ecosystem, but a fair number of them are assigned to multiple ecosystems by means of proportions. Wolverine (Jerv), for example, is assigned with 25\% to forest and and 75\% to mountain. This basic information could easily be displayed on each indicator's page on naturindeks.no. 

The relevant information is found in the assembled indicator data under $indicators: 

```{r, eval = TRUE}

i <- "jerv"
indexData <- readRDS(paste0("data/", i, "_assemebled.rds"))

str(indexData$indicators)

```

Any ecosystem type relevant to a specific indicator appears as a separate column in this dataframe, and contains a value representing the \% fidelity to that ecosystem type. 

Using separately stored information on available ecosystem types, we can assemble this data for all of our example indicators: 

```{r, eval = TRUE}
# Load ecosystem info
EcoSysInfo <- readRDS("data/EcosystemInfo.rds")

# Indicator list
indicator <- c("Dikesoldogg",
               "Jerv",
               "Elg",
               "Lomvi",
               "Havørn",
               "Lange")

# Assemble fidelity data
fidData <- data.frame()

for(i in 1:length(indicator)){
  
  indexData <- readRDS(paste0("data/", indicator[i], "_assemebled.rds"))
  
  ColIdx <- which(names(indexData$indicators) %in% EcoSysInfo$ecosystem)

  fidDataI <- data.frame(
    indicator = indicator[i],
    ecosystem = names(indexData$indicators)[ColIdx],
    fidelity = unname(as.numeric(indexData$indicators[,ColIdx])))
  
  fidData <- rbind(fidData, fidDataI)
}

```

This gives us a dataframe with all indicators and their fidelity to different ecosystems: 
```{r, eval = TRUE}
print(fidData)
```

Before plotting, we match the integer ecosystem IDs to make sure the colour mapping works correctly: 

```{r, eval = TRUE}
fidData <- merge(fidData, EcoSysInfo, all.x = TRUE)
```

Next, we'll visualize this information for each indicator by means of pie charts. 

```{r, eval = TRUE}
suppressWarnings(library(ggplot2))
suppressWarnings(library(tidyverse))

EcoSys_cols <- NIviz_colours$EcoSys_cols[1:11]
names(EcoSys_cols) <- EcoSysInfo$ecosystem

for(i in 1:length(indicator)){
  
  sub_fidData <- subset(fidData, indicator == indicator[i])
  
  print(
    ggplot(sub_fidData, aes(x = "", y = fidelity, fill = fct_inorder(ecosystem))) +
    ggtitle(indicator[i]) + 
    geom_bar(stat = "identity") +
    geom_text(aes(label = paste0(fidelity, "%")),
              position = position_stack(vjust = 0.5)) +
    coord_polar(theta = "y") +
    scale_fill_manual(name = "Ecosystem", values = EcoSys_cols[which(names(EcoSys_cols) %in% sub_fidData$ecosystem)]) + 
    theme_void()
  )

}

```
Depending on how large these pie charts should appear on the website in the end, it may be necessary to move the percentage labels outside the pies. Doing that is a bit more cumbersome, but works with the code below unless the indicator belongs 100 \% to one dataset. When that is the case, the solution below does not print the percentage label at all (see indicator Lange) and I have not figured out how to fix that yet. 


```{r, eval = TRUE}
for(i in 1:length(indicator)){
  
  sub_fidData <- subset(fidData, indicator == indicator[i])
  
  posData <- sub_fidData %>% 
  mutate(csum = rev(cumsum(rev(fidelity))), 
         pos = fidelity/2 + lead(csum, 1),
         pos = if_else(is.na(pos), fidelity/2, pos))
  
  print(
  ggplot(sub_fidData, aes(x = "", y = fidelity, fill = ecosystem)) +
    ggtitle(indicator[i]) + 
    geom_bar(stat = "identity") +
    coord_polar(theta = "y") +
    #scale_fill_manual(values = EcoSys_cols) + 
    scale_fill_manual(values = EcoSys_cols[which(names(EcoSys_cols) %in% sub_fidData$ecosystem)]) + 
    scale_y_continuous(breaks = posData$pos, labels = paste0(sub_fidData$fidelity, "%")) +
    theme(axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_text(size = 15), 
          panel.background = element_rect(fill = "white"))
  )

}



```



